---
output: pdf_document
header-includes:
- \usepackage{orcidlink}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{float}
- \floatplacement{figure}{H}
---

\captionsetup[table]{labelformat=empty}


# Supplementary material

## Brain and body size measurements

Brain size was measured as the weight of the fixed brain for each specimen. First, all collected individuals were kept in cold and sent to the laboratory where they were identified by expert taxonomists and anaesthetised in cold. Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation. Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies. By doing this, we ensured that our brain measurements mainly accounted for neural tissue. We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues. Then, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis®. Finally, we calculated the average weight for the species with more than one specimen. Because weighing bee brains is a highly error-prone task, we excluded values that were under or above 1.5 times the interquartile range for each species.

We considered as a proxy of body size the average intertegular span (ITS) per species. ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight (Cane et al., 2006; Kendall et al., 2019). All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).

Cane, J. H., Minckley, R. L., Kervin, L. J., Roulston, T. A. H., & Williams, N. M. (2006). Complex responses within a desert bee guild (Hymenoptera: Apiformes) to urban habitat fragmentation. _Ecological applications_, _16_(2), 632-644.

Kendall, L. K., Rader, R., Gagic, V., Cariveau, D. P., Albrecht, M., Baldock, K. C., ... & Bartomeus, I. (2019). Pollinator size and its consequences: Robust estimates of body size in pollinating insects. _Ecology and Evolution_, _9_(4), 1702-1714.

\clearpage

```{r Table S1, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}

#Create table of occurrences per habitat type (USA)
#Load libraries
library(tidyverse)
library(kableExtra)

#Load extracted data
pref <- read_csv("../Data/Usa_data/land_cover_usa.csv.gz") %>% 
dplyr::select(species, cover.names) %>% 
mutate_if(is.character,as.factor) 

#Create table with original and new cover categories
pref %>% 
rename(Cover_names = cover.names) %>% 
select(Cover_names) %>% 
group_by(Cover_names) %>% 
summarise(Occurrences = length(Cover_names)) %>% 
mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
       Discarded = "Barren Land", 
       Agricultural = "Cultivated Crops", 
       Natural = "Deciduous Forest", 
       Urban = "Developed, High Intensity", 
       Urban = "Developed, Medium Intensity",
       Urban = "Developed, Low Intensity",
       Urban = "Developed, Open Space", 
       Natural = "Emergent Herbaceous Wetlands", 
       Natural = "Evergreen Forest", 
       Agricultural = "Hay/Pasture", 
       Natural = "Herbaceous", 
       Natural = "Mixed Forest", 
       Discarded = "Open Water", 
       Natural = "Shrub/Scrub", 
       Natural = "Woody Wetlands")) %>% 
arrange(desc(Occurrences)) %>%  
kbl(col.names = c("Original NLCD Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S1.} Original downloaded cover classes for United States with the respective assigned cover classes and number of bee occurrences found on each of them.",
booktabs = T, align = c("l", "c", "l"), linesep = "") %>% 
kable_styling(font_size = 12, latex_options = c("hold_position", "striped")) %>% 
row_spec(0,bold=TRUE)

```

\clearpage

```{r Table S2, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center"}

#Create table of occurrences per habitat type (USA)
#Load libraries
library(tidyverse)
library(kableExtra)

#Load extracted data
pref <- read_csv("../Data/Europe_data/land_cover_europe.csv.gz") %>% 
    dplyr::select(Species, Cover_names) %>% 
    mutate_if(is.character,as.factor) 

#Create table with original and new cover categories
pref %>% 
select(Cover_names) %>% 
group_by(Cover_names) %>% 
summarise(Occurrences = length(Cover_names)) %>% 
mutate(Cover_names1 = fct_recode(as.factor(Cover_names),
      Urban = "Airports", #1
      Discarded = "Bare rocks", #2
      Natural = "Beaches, dunes, sands", #3
      Natural = "Broad-leaved forest", #4
      Natural = "Coastal lagoons", #5
      Agricultural = "Complex cultivation patterns", #6
      Natural = "Coniferous forest", #7
      Urban = "Construction sites", #8
      Urban = "Continuous urban fabric", #9
      Urban = "Discontinuous urban fabric", #10
      Discarded = "Dump sites", #11
      Discarded = "Estuaries", #12
      Agricultural = "Fruit trees and berry plantations", #13
      Urban = "Green urban areas", #14)
      Urban = "Industrial or commercial units", #15
      Natural = "Inland marshes", #16
      Natural = "Intertidal flats", #17
      Agricultural = "Land principally occupied by agriculture, with significant areas of natural vegetation", #18
      Discarded = "Mineral extraction sites", #19
      Natural = "Mixed forest", #20
      Natural = "Moors and heathland", #21
      Agricultural = "Natural grasslands", #22
      Agricultural = "Non-irrigated arable land", #23
      Agricultural = "Pastures", #24
      Natural = "Peat bogs", #25
      Urban = "Port areas", #26) 
      Discarded = "Road and rail networks and associated land", #27
      Natural = "Salt marshes", #28
      Discarded = "Sea and ocean", #29
      Natural = "Sparsely vegetated areas", #30
      Urban = "Sport and leisure facilities", #31
      Natural = "Transitional woodland-shrub", #32
      Agricultural = "Vineyards", #33
      Discarded = "Water bodies", #34
      Discarded = "Water courses" #35  
)) %>% 
mutate(Cover_names = fct_recode(as.factor(Cover_names),
      "Land principally occupied by agriculture" = "Land principally occupied by agriculture, with significant areas of natural vegetation")) %>% 
arrange(desc(Occurrences)) %>%  
kbl(col.names = c("Original CLC Cover class", "Occurrences", "Assigned cover class"), caption = "\\textbf{Table S2.} Original downloaded cover classes for Europe with the respective assigned cover classes and number of bee occurrences found on each of them.",booktabs = T,
align = c("l", "c", "l"), linesep = "") %>% 
kable_styling(font_size = 12,latex_options = c("hold_position", "striped")) %>% 
row_spec(0,bold=TRUE)

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=7, fig.cap="\\textbf{Figure S1.} Association between the log-transformed values of brain and body size (N = 116). The different raw values or points are coloured by the subfamily taxonomic rank.", fig.pos="p", out.width="80%"}

#Plot encephalization
#Load libraries
library(stringr)
library(readr)
library(ggplot2)
library(dplyr)
library(elementalist) #devtools::install_github("teunbrand/elementalist")

#Read data
m = read_csv("../Data/Processing/residuals_model.csv")
r = read_csv("../Data/Processing/residuals_data.csv")
d = read_csv("../Data/Processing/taxonomic_rank.csv") %>% 
mutate(Species = str_replace_all(query, " ", "_")) %>% 
select(!c(query, db))

#Add taxonomic rank to dataset
r = left_join(r,d)

#Add ggplot
ggplot(m, aes(x = log_it, y = estimate__)) +
geom_line(size=1) +
geom_point(data = r, aes(log_it, log_brain_weight, color = subfamily), size=3.5) + 
theme_bw() +
ylab("log(Brain size)") +
xlab("log(Body size)") +
theme(legend.position = c(.95, .60),
legend.justification = c("right", "top"),
legend.box.just = "right",
legend.background = element_rect_round(radius = unit(0.2, "snpc"),fill = "grey85"),
legend.key = element_rect(fill = "grey85")) +
guides(colour=guide_legend(title="Subfamily")) +
theme(axis.title = element_text(face = "bold"))

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure S2.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species. A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}

#Plot maps with density of bee occurrences

#Europe----
#Load libraries
library(giscoR) #map of europe
library(ggplot2)
library(dplyr)
library(readr)
library(sf)
library(patchwork)
library(rnaturalearth)

#Load europe map
Europe <- gisco_get_countries(region = "Europe")
#Load countries filtered
euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Load data with records filtered
all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")

#Plot Europe map with counties of interest highlighted
p1 = ggplot(Europe) +
geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white"))

p2 = ggplot(Europe) +
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15) 

#Insert general plot within specific plot
panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE) 

#USA-----
#Load libraries

#Load data
usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Load maps
canada <- ne_states(country = "canada", returnclass = "sf") %>%
st_as_sf(crs = 4326) 

usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
st_as_sf(crs = 4326)

p1 = ggplot(canada) + 
geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white")) +
geom_map(data = ny_metropolitan_1, 
map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
size = 0.1) +
geom_map(data = usa_states_1, map = usa_states_1,
aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
size = 0.1) +
coord_sf(xlim = c(-86, -62), ylim = c(25, 48)) 

p2 = ggplot(canada) + 
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))

#Insert general plot within specific plot
panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE) 

library(patchwork)
panel2  + panel1



```



\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure S3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban). The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}

#Histogram of habitat preferences
#Load libraries
library(tidyverse) 

#Load preferences
preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv") 

#Merge datasets
preferences = bind_rows(preferences_europe, preferences_usa)

#Convert to long to model everything at the same time
long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))

#Relevel for plotting
long_data = long_data %>% 
mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 

labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")

long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))

long_data = long_data %>% 
mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))

long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))

ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
    xlab("Habitat preference") +
    ylab("Frequency") +
    facet_wrap(~ Habitat) +
    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    theme_bw() +   
    theme(axis.title = element_text(face = "bold", size = 14),
    panel.border=element_rect(size=1),
    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
    panel.grid.minor=element_blank())+
    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))




```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure S4}. Association between relative brain size (A and D), brain weight (B and E) and intertegular distance (C and F) with habitat preference separated by habitat type (i.e., natural, agricultural and urban) and geographical region (USA in the upper panel and Europe in the lower one). The shaded and coloured areas by habitat type represent 95\\% credible intervals.", fig.pos="p", out.width="100%"}

plot_all = readRDS("../Data/Processing/plot_all.rds")

plot_all

```



\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure SX}. ", fig.pos="p", out.width="100%"}

#To answer one of the issues raised by a reviewer
#We explore how the bee's diet (oligo/polylectic)
#is associated with habitat preference (1) and brain size (2)

#The data of the diet comes from the paper of Ferran
#but because is missing for some species
#I have look the diets for missing ones (~20spp)

#Load libraries
library(dplyr)
library(stringr)
library(ggplot2)
library(patchwork)
library(readr)

# (1) Lecticity ~ habitat preference ----
#First we explore this association with habitat preference
#Load preferences data
preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
#Bind rows
pref = bind_rows(preferences_usa, preferences_eu)

#Average preferences between USA and Europe
preferences = pref %>% 
group_by(Species) %>%
summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
ungroup()

#Read lecticity
lecticity = read_csv("../Data/Bees_Lecticity.csv") 
#Delete underscore from tree
lecticity$species = str_replace(lecticity$species, "_", " ")
lecticity = lecticity %>% rename(Species = species)

#Bind datasets
d = left_join(preferences, lecticity)

#Check species with lacking info
#Agapostemon virescens
#https://lopezuribelab.com/halictidae/agapostemon-virescens/#:~:text=Agapostemon%20virescens%20is%20a%20polylectic,coneflowers%2C%20asters%2C%20and%20goldenrod.
#As the other Agapostemon, they tend to visit many unrelated plant species
#but they tend to prefer Asteraceae flowers
#Polylectic
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Agapostemon virescens", "Polylectic"))
#Andrena angustior
#http://dx.doi.org/10.1016/j.biocon.2017.09.009
#Polylectic
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena angustior", "Polylectic"))
#Andrena hirticincta
#https://www.sharpeatmanguides.com/andrena-mining-bees
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena hirticincta", "Oligolectic"))
#Andrena milwaukeensis
#https://www.discoverlife.org/20/q?search=Andrena+milwaukeensis
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena milwaukeensis", "Polylectic"))
#Andrena pilipes
#https://www.wildbienen.de/eb-apili.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena pilipes", "Polylectic"))
#Andrena vicina
#https://www.discoverlife.org/20/q?search=Andrena+vicina
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena vicina", "Polylectic"))
#Andrena wilkella
##http://dx.doi.org/10.1016/j.biocon.2017.09.009
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena wilkella", "Oligolectic"))

#Augochlorella aurata
#https://www.beesofcanada.com/species/augochlorella-aurata
#https://www.discoverlife.org/20/q?search=Augochlorella+aurata
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Augochlorella aurata", "Polylectic"))

#Augochloropsis metallica
#https://www.gbif.org/species/1353091
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Augochloropsis metallica", "Polylectic"))

#Bombus griseocollis
#https://doi.org/10.1002/ecy.2697
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Bombus griseocollis", "Polylectic"))

#Calliopsis andreniformis
#https://www.sharpeatmanguides.com/calliopsis-andreniformis
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Calliopsis andreniformis", "Polylectic"))

#Ceratina strenua
#https://www.beesofcanada.com/species/ceratina-strenua
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Ceratina strenua", "Polylectic"))

#Colletes thoracicus
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Colletes thoracicus", "Polylectic"))

#Osmia atriventris
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia atriventris", "Polylectic"))

#Osmia lignaria
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia lignaria", "Polylectic"))

#Osmia pumila
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia pumila", "Polylectic"))

#Osmia cornifrons
#http://dx.doi.org/10.1098/rsos.200225
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia cornifrons", "Polylectic"))

#Lasioglossum malachurum
#https://www.wildbienen.de/eb-lmala.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Lasioglossum malachurum", "Polylectic"))

#Megachile texana
#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))

#Megachile texana
#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))


#Megachile willughbiella
#https://www.wildbienen.de/eb-mwill.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))


#Natural
p1_1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Habitat Preference")+
xlab(NULL)+
ggtitle("Natural")+
theme_bw() 


#Agricultural
p1_2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab(NULL)+
xlab("Diet")+
ggtitle("Agricultural")+
theme_bw()


#Urban
p1_3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab(NULL)+
xlab(NULL)+
ggtitle("Urban")+
theme_bw()

panel1 = p1_1+p1_2+p1_3 


# (2) Lecticity ~ brain size ----
#Load brain data
brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")

d1 = left_join(brain_weight, d) %>% 
filter(!is.na(Lecticity))

p2_1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Relative brain size")+
xlab("Diet")+
theme_bw()


p2_2 = ggplot(d1, aes(x = Lecticity, y = Brain.weight)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Absolute brain size")+
xlab("Diet")+
theme_bw()

panel2 =  p2_2 + p2_1


library(cowplot)

plot_grid(plot_grid(panel1, labels = c('A'))+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)),

plot_grid(panel2, labels = c('B'))+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)), nrow = 2)


```
